Fish size response to climate warming

### 2017-10-16 draft.Rnw
draft; close #1

Increasing temperatures under climate change are thought to affect individual physiology of ectotherms through increases in metabolic demands, leading to decreases in body-size with concomitant effects on species ecology. Although intuitively appealing, the magnitude, as well as when and where such changes are expected remains controversial. We developped a general trait-based model of carbon and oxygen budgets to determine the relative influence of thermal sensitivity, oxygen limitation, and environmental conditions on ectotherm body size. We use this model to contrast the response of size-at-maturation and asymptotic size in small and large fish with traits favoring fast growth versus reduced mortality. The immediate (plastic) response of size is govered by evolved maturation reaction norms that are largely independent of traits. The evolutionary response, however, is more pronounced in active species, suggesting a potentially stronger reponse to climate drivers. We suggest that this model provides a mechanistic underpinning that ties together various strands of thought and observation about ecological consequences of increasing temperatures.

@@

In the constrained activity model, ectotherms adjust the relative amounts of time ($\tau$) spent foraging and resting/hiding to optimise the net energy/mass gain $P_C$ relative to natural mortality. Since both energy gain and loss are sensitive to temperature and oxygen limitations, both the activity level and the net energy gain will be subject to these environmental constraints. Their interplay thus determines available energy for growth, reproduction, and, ultimately, organisms final size.

The model is written in terms of carbon (or energy) and oxygen balance equations:

\begin{align}
f_C &= \frac{\tau }{\tau  + \frac{h c_T w^{q-p}}{\gamma\Theta} } \label{eq:f} \\
P_C &= S_C - D_C \\
  &=(1-\beta-\phi)f_C h c_T w^q  - c_T k w^n - \tau \delta c_T k w  \\
P_{0_2} &= S_{O_2} - D_{O_2} \\
        &= f_{O_2}w^n - \omega \left( \beta f_C h c_T w^q + c_T k w^n + \tau \delta c_T k w \right)
\end{align}

where $f$ is the feeding level ([0,1]) at weight $w$, determined by the fraction of time spent foraging (or proportion of maximum attack rate) $\tau$, consumption rate $\gamma w^p \Theta$ (search rate $\gamma w^p$ times prey availability $\Theta$) and maximum consumption $h w^q$. In the following, we will refer to $\tau$ as the activity fraction for sake of generality. Maximum consumption, determined by digestive (enzymatic) processes (\cite{jeschke_predator_2002, sentis_parsing_2013}), is assumed to scale with temperature as $c_T = e^{E_a(T-T_0)/kTT_0}$. Available carbon $P_C$ is determined by supply ($S_C$) from prey consumption ($f_C h c_T w^q$), with $\beta$ a loss due to specific dynamic action (SDA, or heat increment; the energy spent absorbing food), and $\phi$ is the fraction of food excreted and egested. Metabolic costs ($D_C$) are those of standard metabolism ($k w^n$), as well as active metabolism (scaled in units of standard metabolism as $\delta k w$, with $\delta$ the factorial aerobic scope (FAS)), with the activity fraction $\tau$ determining the fraction of time that the active metabolism cost applies. Note that active metabolism scales directly with $w$, owing to muscular demands scaling approxiuamtely isometrically (CITE Glazier, Brett).

The oxygen budget $P_{0_2}$, determines physiological limits to aerobic activity. Demand ($D_{O_2}$) is the sum of oxygen used for SDA (i.e., the conversion of food into energy and growth - $\beta f_C h c_T w^q$) and ``active metabolism", the sum of standard (or resting) metabolism and the metabolic costs of foraging ($[1+\tau \delta] c_T k w^n$), with $\omega$ determining amount of oxygen required per unit of metabolised carbon. Metabolic scope is the difference between $S_{O_2}/w^n=f_{O_2}$, the amount of oxygen supplied per unit weight, and oxygen demand due to standard metabolism ($\omega c_T k w^n$).

The maximum oxygen consumption---usually assumed to occur at the maximum metabolic rate (MMR)---is the oxygen consumption during maximal activity level that can be sustained over some time. This level is determined by oxygen delivery to organs and muscles. At constant temperature $T$, oxygen supply is a function of ambient oxygen and is assumed to follow a saturating function \cite[e.g.,][]{lefrancois_influence_2003}. We specify $P_{50}$ as the point where the MMR (or oxygen supply) has dropped by 50\% relative to the saturation level $l$ \footnote{Could also use the same functional form as the consumption of food for simplicity: $f_{O_2} = \Gamma \Theta_{O_2}/\Gamma \Theta_{O_2} + h_{O_2} = \Theta_{O_2}/(\Theta_{O_2} + P_{50}).$}. Oxygen supply, and hence the MMR, are usually found to be temperature dependent \cite{lefevre_are_2016}. We follow \cite{gnauck2013freshwater, lefrancois_influence_2003} and use a dome shaped function that peaks at the optimal temperature (optimal in the sense of maximising the MMR). Thus:

\begin{align}
f_{O_2}&=\lambda(1-e^{O_2\log(0.5)/P50}),\\
\lambda&=\zeta(\frac{T_{\text{max}}-T}{T_{\text{max}}-T_{\text{opt}}})^\eta \times \exp(-\eta\frac{T_{\text{max}}-T}{T_{\text{max}}-T_{\text{opt}}}),
\end{align}

with $f_{O_2}$ the oxygen supply (MMR), $T_{\text{max}}$ the lethal temperature for the species, $T_{\text{opt}}$ the temperature at which the MMR is optimised; $\eta$ determines the width of the dome-shape, and $\zeta$ its height (Figure \ref{fig:O2_fig}). Note that we can emulate a MMR that increases up to the lethal temperature by setting the optimal temperature for MMR close to the maximum temperature.

@@

We now assume that organisms will adjust their activity level to maximise available energy relative to mortality $M$: $\tau_{\text{opt}}$ = $\max_\tau(F)$ = $\max_\tau(P_C/M)$, with $F$ a measure of fitness. We further assume here that mortality scales with activity level and weight as $w^{q-1}$ (\cite{andersen_how_2009, hartvig_food_2011}), such that $M_{\tau,w} = (\rho+\eta\tau) w^{q-1}$, where $ \rho$ is mortality at mass $w=1$ and $\tau=0$, that is, with no activity beyond that covered by standrad metabolism.  $M$ is then the effective mortality at realised levels of activity $\tau$ at weight $w$.

@@

On ecological time-scales, temperature affects size via its effets on the energy budget and the investment of available (surplus) energy into reproduction and growth. The change in allocation to reproduction with size and age in variable environmental conditions is described by the maturation reaction norm, which describes the probability of maturing at a certain age under different growth conditions (\cite[e.g., ][]{dieckmann_probabilistic_2007}). We use a slightly different definition here since our model is not probabilitic at the population level: we define the reaction norm as the mid-point of a logistic allocation function that determines investment in reproduction as a function of age and size. The allocation is parametrised as $\phi(z,c) = 1/(1+\exp(-(c*z)))$, where $z = (w-w^*)\times \text{cos}(\text{atan}(b))-a\times \text{sin}(\text{atan}(b))$ rotates the coordiante system about the slope $b$ of the reaction norm, $a$ is the age, $w$ is the mass, $w^*$ is the intercept of the reaction norm, and $c$ determines how rapidly energy allcoation shifts from somatic growth to reproduction.

The slope of maturation reaction norms is evolutionarily determined by the strength of the co-variation between growth and mortality for a given population in a given environment (\cite{marty_impact_2011}), with strongly positive co-variation leading to the relatively flat reaction norms observed for most populations of fish. We do not explicitly model these interactions here (aside from the dependence of mortality on $\tau$) as we are interested in parsing out the effect of temperature, and we assume that reaction norms evolved over relatively stable temperature regimes in the past. In other words, we assume that temperature was not a key driver of covariations between growth and mortality. In that scenario, good growth conditions (e.g., from increased prey density) probably alter the baseline mortality $\rho$ and the risk of foraging $\eta$ (e.g., by attracting predators), independently of temperature influences. Consequently, we assume that the evolved slope of the reaction norm is a trait for a particular species or population. Since most reaction norms are relatively flat, we simply assume a flat reaction norm.

The maturation reaction norm is subjected to selective pressure when through a gradual change in temperature, mean environmental conditions change and previously evolved reaction norms become sub-optimal. To investigate how size will change over evolutionary time-scales, we calculate the expected life-time reproduction, $R0$ as temperature changes from a reference temperature $T_ref$. We assumed that the covariation between mortality and growth remains similar as temperatures change (i.e., temperature remains independent of covariation between growth and mortality, and ), and we optimised $R0$ with respect to $w*$ at each temperature. From this, we calculated the the expected rate of change (i.e., the selection response) of $w*$ with resepct to temperature as $S = \sigma_w^* h^2 R0'/R0$, where $R0' = dR0/dT$, where we take $h$ to be about 0.2 and we assume a coefficient of variation about $w$ of 0.2.

@@

We explored ecological and evolutionary impacts along two ecological trait axes. The first axis aims to contrast species along a gradient of life-history that, at the one end, maximises production (energy aquisition) at the cost of increased metabolism and mortality, and at the opposing end minimises mortality and metabolic costs at the expense of production. This axis leads to an approximately constant ratio of production to mortality, and corresponds to a line of equal size in the life-history space proposed by \cite{charnov_evolutionary_2013}. In other words, this axis contrasts species of similar size with defensive/sluggish versus active  life-histories.

To implement this axis, we use the result that species with a more active, production oriented life-history (e.g., pelagic species) have a higher standard metabolism and lower weight scaling of metabolic costs \cite{killen_intraspecific_2010}. We assume that that higher standard metabolism is due to increased digestive capacity (i.e., is used for gut maintenance CITE). In practice, we assume that approximately 50\% of the standard metabolic cost is due to gut maintenance, such that a doubeling of the maximum ingestion leads to a 50\% increase in standard metabolic cost. We further assume that such active species have a less effective refuge from predators and therefore have a higher constant mortality, but lower mortality related to activity. Exact parameter values are given in Table \ref{tab:parameters}.

The second trait axis aims to be orthogonal to the first and contrast changes in size for species of large and small absolute size, but with similar investment in production or mortality avoidance (i.e., with the same value on the first trait axis).  Inter-specific meta-analyses often show that metabolism scales with mass to a power of roughly 3/4 (\cite[e.g., ][]{brown_toward_2004}). This contrasts with inter-specific weight scaling, which is often found to be higher than 3/4 (\cite{killen_little_2007,ohlberger_intraspecific_2012}). We thus assume a trait axis based on metabolic cost per unit weight, which naturally leads to differences in absolute size in our model.

@@

Slope & $b$ & 0  & 0 & 0 & 0\\
Reaction & $c$ & 0.5 & 0.1 & 0.5 & 0.1 \\

\addlinespace
\multicolumn{6}{c}{\textbf{Oxygen budget}} \\
Critical $O_2$ & $P_{crit}$ ($mg\cdot L^{-1}$) & 2 & 2& 2& 2\\
Disolved $O_2$ at $0.5\times f_{max}(O_2)$  & $P50$ ($mg\cdot L^{-1}$) & 4  & 4  & 4  & 4 \\
Doming for $O_2$ supply & $\eta$ & 3 & 3 & 3 & 3\\
Level of $O_2$ supply  & $\zeta$ ($g\cdot y^{-1}$) &  1 & 1 & 2 & 2\\

@@

@@

int_m <- 1000
int_t <- 1000
int_temp <- 30

# Sluggish small
ss <- list(c=0.3,
           beta=0.15,
           delta=4,
           Ea=0.52, 
           gamma=60, 
           h=30, 
           k=1,
           lO=0.5, 
           m=10, 
           n=0.8, 
           O2crit=2, 
           omega=0.25, 
           p=0.8,
           P50=4.2, 
           phi=0.25, 
           q=0.8,
           shape=3,
           temp=seq(5,26,l=int_temp),
           Topt=20, 
           temp_ref=15,
           slope=-0.3,
           tmax=10,
           tr=0.5,
           M=0.1,
           v=6,
           lm=int_m,
           nu=0.8-1,
           dt=int_t,
           min_m=-2,
           max_m= 3
)

@@

fl <- fs
fl$n <- 0.75
fl$k <- 1
fl$tr <- 0.1
fl$tmax <- 20
fl$max_m <- 4

sl <- ss
sl$n <- 0.8
sl$k <- 0.7
sl$tmax <- 50
sl$max_m <- 4
sl$tr <- 0.1

@@

<<activity,fig.cap='With increasing temperature, most organisms will increase activity $\\tau$ to offset increased metabolic costs. Oxygen sets metabolic limits at extreme temperatures. Responses are shown for arbitrary M- and P-strategy (left and right columns, respectively), for small and large species (top and bottom rows, respectively).'>>=

@@

<<feeding,fig.cap='Feeding level, available energy and mortality are impacted by increased activity and the metabolic response to temperature. Responses are shown for arbitrary M- and P-strategy (left and right columns, respectively), for small and large species (top and bottom rows, respectively). Energy is plotted relative to maximum over all temperatures.'>>=

@@

<< norms, fig.cap='On ecological time-scales, temperature affects size via evolved maturation-reaction norms (\\cite{marty_impact_2011}). Most reaction norms found in nature have a negative slope, meaning for faster growth, maturation occurs at a larger size. Increasing temparture (purple to yellow growth curves) modifies growth, and maturation age changes according to the reaction norm (bottom left panel), whereas asymptotic size is affected by changes in absolute energy available for growth.'>>=

@@

<<evo, fig.cap='On evolutionary time-scales, selection moves the reaction norm. Size-at-maturation reacts strongly to temperature changes in the model (top left panel). Selection (S) induced changes in size are slow for M strategists, fast for P strategists. (top right panel). In absolute terms, these changes are slow for both small (bottom left panel) and large fish (bottom right panel)',fig.height=6>>=

@@

@@

Life history theory suggests that the adult weight $W_{\infty}$ is found at the size at which all energy is allocated to reproduction, which optimally occurs at $\frac{\partial P_C/M}{\partial w}=1$ in the case of size-dependent mortality (provided concavity with respect to mass of this quantity). Note that $W_{\infty}$ depends on temperature only through $\tau$ and changes in $P_C$ as temperature changes metabolic and enzymatic rates. We thus impose no direct constraint on $W_{\infty}$.

>>>>

Increasing temperatures under climate change are thought to affect individual physiology of ectotherms through increases in metabolic demands, leading to decreases in body-size with concomitant effects on species ecology. Although intuitively appealing, the magnitude, as well as when and where such changes are expected remains controversial. We developed a general trait-based model of carbon and oxygen budgets to determine the relative influence of thermal sensitivity, oxygen limitation, and environmental conditions on ectotherm body size. We use this model to contrast the response of size-at-maturation and asymptotic size in small and large fish with traits favoring fast growth versus reduced mortality. The immediate (plastic) response of size is governed by evolved maturation reaction norms that are largely independent of traits. The evolutionary response, however, is more pronounced in active species, suggesting a potentially stronger response to climate drivers. We suggest that this model provides a mechanistic underpinning that ties together various strands of thought and observation about ecological consequences of increasing temperatures.

@@

In the constrained activity model, ectotherms adjust the relative amounts of time ($\tau$) spent foraging and resting/hiding to optimise the net energy/mass gain $P$ relative to natural mortality. Since both energy gain and loss are sensitive to temperature and oxygen limitations, both the activity level and the net energy gain will be subject to these environmental constraints. Their interplay thus determines available energy for growth, reproduction, and, ultimately, organisms final size.

The model is written in terms of weight $w$ and oxygen balance equations at temperature $T$:

\begin{align}
f(w,T) &= \frac{\tau }{\tau  + \frac{h c_T w^{q-p}}{\gamma\Theta} } \label{eq:f} \\
P(w,T) &= S(w,T) - D(w,T) \\
  &=(1-\beta-\phi)f_C h c_T w^q  - c_T k w^n - \tau \delta c_T k w  \\
P(w,T)_{0_2} &= S(w,T)_{O_2} - D(w,T)_{O_2} \\
        &= f_{O_2}(t)w^n - \omega \left( \beta f(w,T) h c_T w^q + c_T k w^n + \tau \delta c_T k w \right)
\end{align}

where $f(w,T)$ is the feeding level ([0,1]) at weight $w$ and temperature $T$, determined by the fraction of time spent foraging (or proportion of maximum attack rate) $\tau$, consumption rate $\gamma w^p \Theta$ (search rate $\gamma w^p$ times prey availability $\Theta$) and maximum consumption $h w^q$. In the following, we will refer to $\tau$ as the activity fraction for sake of generality. Maximum consumption, determined by digestive (enzymatic) processes (\cite{jeschke_predator_2002, sentis_parsing_2013}), is assumed to scale with temperature as $c_T = e^{E_a(T-T_0)/kTT_0}$. Available carbon $P(w,T)$ is determined by supply ($S(w,T)$) from prey consumption ($f_C h c_T w^q$), with $\beta$ a loss due to specific dynamic action (SDA, or heat increment; the energy spent absorbing food), and $\phi$ is the fraction of food excreted and egested. Metabolic demands ($D(w,T)$) are those of standard metabolism ($k w^n$), as well as active metabolism, scaled in units of standard metabolism as $\delta k w$, with $\delta$ the factorial aerobic scope (FAS)). The activity fraction $\tau$ determines the fraction of time that the active metabolism cost applies. Note that active metabolism scales directly with $w$, owing to muscular demands scaling approximately isometrically with weight (\cite{glazier_activity_2009,brett1965relation}). In the following we drop dependencies on weight and temperature in the notation for sake of readability.

The oxygen budget $P_{0_2}$, determines physiological limits to aerobic activity. Demand ($D_{O_2}$) is the sum of oxygen used for SDA (i.e., the conversion of food into energy and growth - $\beta f h c_T w^q$) and ``active metabolism", the sum of standard (or resting) metabolism ($c_T k w^n$) and the metabolic costs of foraging ($\tau \delta c_T k w$), with $\omega$ determining amount of oxygen required per unit of metabolised carbon. Metabolic scope is the difference between $S_{O_2}/w^n=f_{O_2}(T)$, the maximum amount of oxygen supplied per unit weight at temperature $T$, and oxygen demand due to standard metabolism ($\omega c_T k w^n$).

The maximum oxygen consumption---usually assumed to occur at the maximum metabolic rate (MMR)---is the oxygen consumption during maximal activity level that can be sustained over some time. At constant temperature $T$, oxygen supply is a function of ambient oxygen and is assumed to follow a saturating function \cite[e.g.,][]{lefrancois_influence_2003}. We specify $P_{50}$ as the point where oxygen supply has dropped by 50\% relative to the saturation level $l$ \footnote{Could also use the same functional form as the consumption of food for simplicity: $f_{O_2} = \Gamma \Theta_{O_2}/\Gamma \Theta_{O_2} + h_{O_2} = \Theta_{O_2}/(\Theta_{O_2} + P_{50}).$}. Oxygen supply, and hence the MMR, are usually found to be temperature dependent \cite{lefevre_are_2016}. We follow \cite{gnauck2013freshwater, lefrancois_influence_2003} and use a flexible dome shaped function:

\begin{align}
f_{O_2}(T)&=\lambda(T)(1-e^{O_2\log(0.5)/P50}),\\
\lambda(T)&=\zeta(\frac{T_{\text{max}}-T}{T_{\text{max}}-T_{\text{opt}}})^\eta \times \exp(-\eta\frac{T_{\text{max}}-T}{T_{\text{max}}-T_{\text{opt}}}),
\end{align}

with $f_{O_2}$ the oxygen supply. Here $\lambda(T)$ specifies the temperature dependency of $O_2$ supply, whereas the second term describes the saturation dependence on ambient $O_2$. To specify $\lambda(T)$, we use $T_{\text{max}}$ the lethal temperature for the species, $T_{\text{opt}}$ the temperature at which oxygen supply is maximised; $\eta$ determines the width of the dome-shape, and $\zeta$ its height (Figure \ref{fig:O2_fig}). Note that we can emulate an oxygen supply (and hence maximum metabolic rate; MMR) that increases up to the lethal temperature by setting the temperature for maximum oxygen delivery close to the lethal temperature (Figure \ref{fig:O2_fig}).

@@

We now assume that fish will adjust their activity level to maximise fitness $F$: $\tau_{\text{opt}}$ = $\max_\tau(F)$ = $\max_\tau(P/M)$, which amounts to maximizing available energy relative to mortality $M$ \cite{gilliam_habitat_1987,sainmont_effective_2015}). We further assume here that mortality scales with activity level and weight as $w^{q-1}$ (\cite{andersen_how_2009, hartvig_food_2011}), such that $M(w) = (\rho+\eta\tau) w^{q-1}$, where $ \rho$ is mortality at mass $w=1$ and $\tau=0$, that is, with no activity beyond that covered by standard metabolism. $M$ is then the effective mortality at realised levels of activity $\tau$ at weight $w$.

@@

@@

We explored ecological and evolutionary impacts along two ecological trait axes. The first axis aims to contrast species along a gradient of life-history that, at the one end, maximises production (energy acquisition) at the cost of increased metabolism and mortality, and at the opposing end minimises mortality and metabolic costs at the expense of production. This axis leads to an approximately constant ratio of production to mortality, and corresponds to a line of equal size in the life-history space proposed by \cite{charnov_evolutionary_2013}. In other words, this axis contrasts species of similar size with defensive/sluggish versus active  life-histories.

To implement this axis, we use the result that species with a more active, production oriented life-history (e.g., pelagic species) have a higher standard metabolism and lower weight scaling of metabolic costs \cite{killen_intraspecific_2010}. We assume that that higher standard metabolism is due to increased digestive capacity (i.e., is used for gut maintenance CITE). In practice, we assume that approximately 50\% of the standard metabolic cost is due to gut maintenance, such that a doubling of the maximum ingestion leads to a 50\% increase in standard metabolic cost. We further assume that such active species have a less effective refuge from predators and therefore have a higher constant mortality, but lower mortality related to activity. Exact parameter values are given in Table \ref{tab:parameters}.

The second trait axis aims to be orthogonal to the first and contrast changes in size with temperature for species of large and small absolute size, but with similar investment in production or mortality avoidance (i.e., with the same value on the first trait axis).  Inter-specific meta-analyses often show that resting metabolism scales with mass to a power of roughly 3/4 (\cite[e.g., ][]{brown_toward_2004}). This contrasts with inter-specific weight scaling, which is often found to be higher than 3/4 (\cite{killen_little_2007,ohlberger_intraspecific_2012}). We thus assume a trait axis based on metabolic cost per unit weight, which naturally leads to differences in absolute size in our model.

@@

Slope & $b$ & 0 (-0.5,0.5)  & 0 (-0.5,0.5) & 0 (-0.5,0.5) & 0 (-0.5,0.5)\\
Reaction & $c$ & 0.5 & 0.1 & 0.5 & 0.1 \\

\addlinespace
\multicolumn{6}{c}{\textbf{Oxygen budget}} \\
Critical $O_2$ & $P_{crit}$ ($mg\cdot L^{-1}$) & 2 & 2& 2& 2\\
Dissolved $O_2$ at $0.5\times f_{max}(O_2)$  & $P50$ ($mg\cdot L^{-1}$) & 4  & 4  & 4  & 4 \\
Doming for $O_2$ supply & $\eta$ & 3 & 3 & 3 & 3\\
Level of $O_2$ supply  & $\zeta$ ($g\cdot y^{-1}$) &  1 & 1 & 2 & 2\\

@@

On ecological time-scales, temperature affects size via its effects on the energy budget and the investment of available (surplus) energy into reproduction and growth. The change in allocation to reproduction with size and age in variable environmental conditions is described by the maturation reaction norm, which is generally defined as the probability of maturing at a certain age under different growth conditions (\cite[e.g., ][]{dieckmann_probabilistic_2007}). We use a slightly different definition here since our model is not probabilistic at the population level: we define the reaction norm as the mid-point of a logistic allocation function that determines investment in reproduction as a function of age and size. The allocation is parametrised as 
$$\phi(z) = 1/(1+\exp(-(c*z))) \text{, where }$$
$$z(t, w_t) = (w_t-w^*)\times \text{cos}(\text{atan}(b))-t\times \text{sin}(\text{atan}(b))$$ 
rotates the coordinate system about the slope $b$ of the reaction norm, $t$ is the age, $w$ is the mass, $w^*$ is the intercept of the reaction norm, and $c$ determines how rapidly energy allocation shifts from somatic growth to reproduction.

The slope of maturation reaction norms is evolutionarily determined by the strength of the co-variation between growth and mortality for a given population in a given environment (\cite{marty_impact_2011}). Strongly positive co-variation leads to the relatively flat reaction norms observed for most populations of fish. We do not explicitly model these interactions here (aside from the dependence of mortality on $\tau$) as we are interested in parsing out the effect of temperature, and we assume that reaction norms evolved over relatively stable temperature regimes in the past. In other words, we assume that temperature was not a key driver of covariation between growth and mortality. In that scenario, good growth conditions (e.g., from increased prey density) probably alter the baseline mortality $\rho$ and the risk of foraging $\eta$ (e.g., by attracting predators). Consequently, we assume that the evolved slope of the reaction norm is a trait for a particular species or population.

The maturation reaction norm is subject to selective pressure when through a gradual change in temperature, mean environmental conditions change and previously evolved reaction norms become sub-optimal. To investigate how size will change over evolutionary time-scales, we calculated the expected life-time reproduction $R0$ with respect to $w^*$ as temperature $T$ changes from a reference temperature $T_ref$:

$$R0 = \int_0^\infty \phi(z(t,w_t,w^*))P(w_t,T)S_{0 \rightarrow t} dt, \text{, where}$$
$$S_{0 \rightarrow t} = \int_0^t \exp(-M(w_t))dt$$

is survival from birth to age $t$. We assumed that the covariation between mortality and growth remains similar as temperatures change (i.e., temperature remains independent of covariation between growth and mortality, and the reaction norm slope does not change substantially), and we optimised $R0$ with respect to $w^*$ at each temperature. From this, we calculated the the expected rate of change (i.e., the selection response) of $w^*$ with respect to temperature as $S = \sigma_w^* h^2 R0'/R0$, where $R0' = dR0/dT$, where we take the heritability term $h^2$ to be about 0.2 and we assumed a coefficient of variation about $w$ of 0.2 for $\sigma_w$.

@@

int_m <- 1000
int_t <- 1000
int_temp <- 50

# Sluggish small
ss <- list(c=0.3,
           beta=0.15,
           delta=4,
           Ea=0.52, 
           gamma=60, 
           h=30, 
           k=1,
           lO=0.5, 
           m=10, 
           n=0.8, 
           O2crit=2, 
           omega=0.25, 
           p=0.8,
           P50=4.2, 
           phi=0.25, 
           q=0.8,
           shape=3,
           temp=seq(5,26,l=int_temp),
           Topt=20, 
           temp_ref=15,
           slope=-.3,
           tmax=15,
           tr=0.5,
           M=0.1,
           v=6,
           lm=int_m,
           nu=0.8-1,
           dt=int_t,
           min_m=-2,
           max_m= 3
)

@@

fl <- fs
fl$k <- 1
fl$tr <- 0.1
fl$tmax <- 20
fl$max_m <- 4

sl <- ss
sl$n <- 0.8
sl$k <- 0.8
sl$tmax <- 50
sl$max_m <- 4
sl$tr <- 0.1

@@

<<activity,fig.cap='With increasing temperature, most organisms will increase activity $\\tau$ to offset increased metabolic costs. Oxygen sets metabolic limits at extreme temperatures. Responses are shown for arbitrary M- and P-strategy (left and right columns, respectively), for small and large species (top and bottom rows, respectively).',fig.align='center'>>=

@@

<<feeding,fig.cap='Feeding level, available energy and mortality are impacted by increased activity and the metabolic response to temperature. Responses are shown for arbitrary M- and P-strategy (left and right columns, respectively), for small and large species (top and bottom rows, respectively). Energy is plotted relative to maximum over all temperatures.',fig.align='center'>>=

@@

<< norms, fig.cap='On ecological time-scales, temperature affects size via evolved maturation-reaction norms (\\cite{marty_impact_2011}). Most reaction norms found in nature have a negative slope, meaning for faster growth, maturation occurs at a larger size. Increasing temparture (purple to yellow growth curves) modifies growth, and maturation age changes according to the reaction norm (bottom left panel), whereas asymptotic size is affected by changes in absolute energy available for growth.',fig.align='center'>>=

@@

<<evo, fig.cap='On evolutionary time-scales, selection moves the reaction norm. Size-at-maturation reacts strongly to temperature changes in the model (top left panel). Selection (S) induced changes in size are slow for M strategists, fast for P strategists (top right panel; note that selection response is normalised by size here for comparison, i.e., cahnges are shown proportional to size at each temperature). In absolute terms, these changes are slow for both small (bottom left panel) and large fish (bottom right panel)',fig.height=6,fig.align='center'>>=

@@

<<goutss>>=

fs_n <- fs
fs_n$slope <- -0.5
fs_p <- fs
fs_p$slope <- 0.5

scss <- list(fs,fs_n,fs_p)
scns <- c(0,-0.5,0.5)
goutss <- parallel::mclapply(scss,plot_data_growth_tO2,mc.cores = 4)
@

<<plot_gouts, fig.cap='Selection response for changes in size with temperature (in mm per year) for populations with positive, flat and negative reaction norm slopes. The slope of the reaction norm does not influence the evolutionary response (assuming the slope stays fixed).',fig.height=3,fig.width=4,fig.align='center'>>=

winfs <- bind_rows(lapply(1:length(goutss), function(x) {
  y <- goutss[[x]][['winfs']]
  y$Scenario <- as.factor(scns[x])
  y}))

wtfss <- winfs %>% group_by(Scenario) %>%
  mutate(length=sc(lw(l)))

ggplot(wtfss) +
  geom_line(aes(x=Temperature, y=G, linetype=Scenario)) +
  theme(panel.background = element_blank(),
        plot.background = element_blank())+
  scale_linetype_discrete('Slope')+
  labs(x = expression("Mean temperature " ( degree*C)),
       y=expression(S~response~mm.y^-1))

@@

Life history theory suggests that the adult weight $W_{\infty}$ is found at the size at which all energy is allocated to reproduction, which optimally occurs at $\frac{\partial P/M}{\partial w}=1$ in the case of size-dependent mortality (provided concavity with respect to mass of this quantity). Note that $W_{\infty}$ depends on temperature only through $\tau$ and changes in $P$ as temperature changes metabolic and enzymatic rates. We thus impose no direct constraint on $W_{\infty}$.
